---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# saturnR

## Overview
This package is a wrapper on saturn traffic modelling software to improve saturn and R workflow. It's functions use SATURN built in XEXES as a backend and allow R users to interact with the most common SATURN files such as UFM matrices and UFS files

Currently you can:

* read and write matrices (both single and multiuser) (MX backend)
* read ufs files and extract the geometry and data (P1XDUMP + SATDB backend) which allows to obtain sf objects.

## Installation


```{r eval=FALSE}
# You can install the the development version from GitHub:
# install.packages("devtools")
devtools::install_github("r-aecom/saturnR")
```


## Intro to Matrix manipulation

Matrix manipulation using the built in tools in Saturn can be cumbersome whith this package you can load the data into R modify it and save it again.

The Package comes with some example files for you to test you can see them with
```{r}
library(saturnR)
saturnR_example()
```

To modify the example Epsom98 matrix first we need to set up out XEXES folder as our back end.

```{r}
set_xexes("C:/Program Files (x86)/Atkins/SATURN/XEXES 11.4.07H MC X7/") # Using a particular version of the backend
```

The we can read the matrix. It only has one user class, 12 zones and 144 rows

```{r}
epsom_mat <- saturnR_example("Epsom98mat.UFM") 

mat <- read_ufm(epsom_mat)

mat

```

If we wanted to anlayse the trips between zone 3 and 4 we could use:

```{r}
selected_trips <- (mat$origin == 3 & mat$destination == 4) | # from 3 to 4
  (mat$origin == 4 & mat$destination == 3) # or from 4 to 3

mat[selected_trips,]

```

Then we can increase them by a fixed ammount.

```{r}
fixed_increase <- 200 # A new development adds 200 trips

mat$trips[selected_trips] <- mat$trips[selected_trips] + 200

mat[selected_trips,]

```

Finally we could save the matrix
```{r eval=FALSE}
write_ufm(mat, "Epsom_new_dev.UFM")
```

