---
title: "Reading UFS data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading UFS data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(saturnR)
library(mapview) # for plotting purposes
```

It is common to want to extract link data from UFS files to analyse it, to produce better looking reports
or to send it to other design teams.

In those cases dumping data from UFS files can become tedius and mistake prone. The typical tools like
batch files tend to be rigid and difficult to addapt.

Reading a simple file with read_ufs will dump all available variables from the UFS. Incomplete ufs file report
warnings at the moment due to miss one or more columns (developing)

```{r, warning=FALSE}
# As always we define our XEXES backend in this case 11.4
set_xexes("C:/Program Files (x86)/Atkins/SATURN/XEXES 11.4.07H MC X7/")

epsom_ufs <- saturnR_example("Epsom98oba.UFS")

ufs <- read_ufs(epsom_ufs, load_geometry = TRUE, selection_mode = "all_links")

head(data.frame(ufs[,1:11])) #prin the first few lines and columns
```

As we are obtaining the coordinates we can now export as a shp file
```{r, eval=FALSE}
sf::st_write(ufs, "epsom.shp")
```

Or see the result on a leaflet map

```{r}
mapview(ufs,zcol = "net_speed") #show the net speed in the map (directions overlapping in this example)
```

